### 2022-02-03 paper.tex
first commit

\documentclass{article}
\usepackage[backend=biber, annotation]{biblatex-chicago}
\usepackage[margin=1in]{geometry}
\addbibresource{paper.bib}

\begin{document}

  \section{Topic that I'm currently considering}

  The current topic that I'm considering is Pakistani music. Although currently
  I'm not very knowledgeable about the subject, I do have a vague idea of what I
  want to study since I am from the region (Pakistan), have a basic
  understanding of its history and also some understanding of the language.
  Currently, I'm thinking of taking a few approaches, Firstly, there is the
  colonial aspect. Since Pakistan was ruled by Britain, I'm considering finding
  music that helps to show how colonialism affected pakistan or how the people
  of the region coped with it. The second aspect is the religious one. In Sufi
  Islam, there are mystical pieces of music called the Qawwalis which are
  somewhat popular. Trying to understand how that was a tool used to cope with
  colonialism could also be illuminating.

  \nocite{*}
  \printbibliography
\end{document}